---
title: "Class10"
author: "Nathaniel Bloom"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Human Breast Cancer Cell Analysuis 

```{r}
wisc.df <- read.csv("https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv")

#Convert to matrix of data we want and exclude some columns (ID, Diagnosis, X)
wisc.data <- as.matrix(wisc.df[ ,3:32 ])
head(wisc.data)

```

How many patient data?
```{r}
nrow(wisc.data)
```
How many cancer and non-cancer?
```{r}
table(wisc.df$diagnosis)
# B   M 
#357 212 
```

How many features with _mean? grep() looks for "pattern"
```{r}
colnames(wisc.data)
#see which column names have "_mean"
output <- grep("_mean", colnames(wisc.data), value = TRUE) #value = TRUE tells what names match 
#count how many there are?
length(output)
# 10 places with "_mean" in the column name 
```

Principal Component Analysis 
```{r}
# check mean and STD b/c we need to see if scaling is needed before continuing 

round( apply(wisc.data,2,sd), 2) #STD 2 = look at column; 1 = use row

#we do need to scale b/c contains values that have large variance 
```
Perform PCA
```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE) #scale = TRUE to scale data 
summary (wisc.pr)
#44% of variance in PC1, PC1 + PC2 = 63% 
#PCA creates a new axis that is better at showing the variance among the data points 
#axis are ordered by most variation, PC1=most variation axis, PC2= second most variation 
#scores based on how much they influence the orientation, position, and length of the axis (PCA axis )
#i.e. score close to zero= not much influence, and further away from zero = more influence in each direction 
#each variable read count x weight = PC score 
```
Plot the PCA 
```{r}
biplot(wisc.pr)
```
See whats in the results 
```{r}
attributes(wisc.pr)
#going to use X for plot b/c its the result of our PCA
```

Scatterplot PCA;  PC1 vs. PC2; colored by expert diagnosis 
```{r}
# Scatter plot observations by components 1 and 2
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis, xlab= "PC1", ylab= "PC2") #benign=black, malignant=red
# big variability in red = many ways for cell to be cancerous, lower variation in black means not many ways to be healthy 
#each dot on the plot is all data for each patient compressed into a single data point by focusing on what is different between each point 
```
Plot PC1 and PC3 
```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[, c(1, 3)], col = wisc.df$diagnosis, 
     xlab = "PC1", ylab = "PC3")
```
CLuster in PC space
```{r}
#can we cluster original data ?
kmeans(wisc.data, centers = 2, nstart = 20)
#Show in New WindowClear OutputExpand/Collapse Output
#K-means clustering with 2 clusters of sizes 438, 131 
```

```{r}
#cluster with hclust 
wisc.hc <- hclust(dist(wisc.data))
plot(wisc.hc) #not a very good plot with raw data 
```
Use PCA results instead for clustering 
```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:3]), method="ward.D2")
plot(wisc.pr.hclust) #wisc.pr$x is the results of PCA
```
Cut the tree to get clusters using cutree()
```{r}
grps3 <- cutree(wisc.pr.hclust, k=3)
table(grps3)
#grps3
#  1   2   3 
#111  92 366 
```
Plot clusters 
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = grps3)
```
Use table() Function to compare $diagnosis vector with our cluster results vector 
```{r}
table(grps3, wisc.df$diagnosis)
#K=3
#grps3   B   M
#    1   0 111
#    2  24  68
#    3 333  33
#K=2
#grps3   B   M
#    1  24 179  #still has misclassified patients false positive and false negative
#    2 333  33

```
Prediction Activity 
```{r}
new <- read.csv("new_samples.csv")
new
```

```{r}
npc <- predict(wisc.pr, newdata=new)
npc
```
Add predicted data to previous PCA plot 
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15, cex=3)#add new data to previous model
text(npc[,1], npc[,2], labels = c(1,2), col = "white") #add lebels to the patients and shows that patient 2 would need to be prioritized to have further checks 
```


