---
title: "Class 12 Structural Bioinformatics pt.1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics 
Inspect the types structures in the main database for 3D biomolecular data 

Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
#read in file 
stats <- read.csv("Data Export Summary.csv", row.names = 1)

#determine percentage of structure as x-ray
total <- sum(stats$Total)
total

percents <- stats$Total/sum(stats$Total) * 100
percents

names(percents) <- rownames(stats)
round(percents, 2)



```


Q2: Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}

percentprotein <- round(sum(stats$Proteins)/sum(stats$Total) * 100, 2) 
percentprotein
```

Q3: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

1289

```{r}

```


