---
title: 'Class 6: R Function'
author: "Nathaniel Bloom"
date: "1/23/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions
## Level 2 Header 
### Level 3 Header 

**File Import** Exploration
Main read function in R is 'read.table()'
Read in txt files 
```{r}
t1 <- read.table("test1 (1).txt", sep = ",", header = TRUE)
```
Or just use 'read.csv()'

```{r}
t1 <- read.csv("test1 (1).txt")
t1
```
```{r}
t2 <- read.table("test2.txt", sep = "$", header= TRUE)
t2
```
```{r}
t3 <- read.table("test3.txt")
t3
```
# Back to Functions
Example 1 
```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
```

Use This Function
```{r}
add(7, 3)
```
Vector as an input 

```{r}
add(c(1, 2, 3, 4))
```

Will add 5 to everything 
```{r}
add(c(1, 2, 3, 4), 5)
```
each vector represents X and Y respectively 
```{r}
add(c(1, 2, 3, 4), c(1,2,3,4))
```

```{r}
#add(c(1, 2, 3, 4), 1, 2, 4)
```
Too many arguments b/c function only allows two arguments 

What is 'range()' function 

```{r}
x <- c(1,1,3,11,12)
max(x)
min(x)
```


```{r}
range(x)
```

Our second function 
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale(x)
```

More Testing 
```{r}
rescale(1:10)
```



```{r}
# how do we get this to work 
rescale( c(1,2,NA,3,10) )
```

```{r}
x<- c(1,2,NA,4,5)
rng<- range(x, na.rm =TRUE)
rng
```

New function to deal with NA 
```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2(c(1,2,NA,4,5))
```

rescale 3 
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 
 print("Hello")
 
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 
 if(plot) {
   print("Dont sing again!")
  plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
 
}
```

```{r}
rescale3(x, plot=FALSE)
```
use library to call a function when needed 
Hands on worksheet 
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, typ="l", ylab="Bfactor")
plotb3(s2.b, typ="l", ylab="Bfactor")
plotb3(s3.b, typ="l", ylab="Bfactor")

```
```{r}
library(bio3d)
s1 <- read.pdb("4AKE")
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
print(s1.chainA)
```


```{r}
s1 <- read.pdb("4AKE") # kinase with drug
```

> Q1 what does 'read.pdb()' return?

Returns list of 8 things of class "pdb", "sse"
```{r}
class(s1)
```

```{r}
aa321(s1$seqres)
```

```{r}
s1$atom
```


```{r}
plotb3(s1.b, typ="l", ylab="Bfactor", col="red")
points(s2.b, typ = "l", col="blue", lwd=2)
points(s3.b, typ = "l", col="purple", lwd=2)
```



```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```


Q3
sse

Sum Of Square Errors
Compute the sum of squared prediction errors (or residual sum of squares) when a linear model is applied to a dataset











```{r}
library(bio3d)
# get the pdb file 
# getpdb gets pdb files from online database and assigns the data to newpdb vector 
getpdb <- function(x){ 
  newpdb <- read.pdb(x) # assign newpdb to the pdb file 
return(newpdb)
}
#check if first function worked 
#print(getpdb("4AKE")) 
# trim function uses the getpdb function to retrieve the pdb file and then uses trim.pdb from Bio3d to only include chain identifier A and atom type CA 
trim <- function(y){# trim the pdb file with function trim
  pdbfortrim <- getpdb(y) #use getpdb to retirve file and assign to pdbfortrim
  trimmedpdb <- trim.pdb(pdbfortrim, chain="A", elety="CA") #use trim.pdb and assign to trimmedpdb
return(trimmedpdb)
}
# check if trim function works 
#print (trim("4AKE"))
# function atomb uses trim function and assigns the trimmed pdb file to gettrim and then retirves only atom b from the trimmed file and assigns it to atombfromtrimmedfile
atomb <- function(x){
   gettrim <- trim(x)
  atombfromtrimmedfile <- gettrim$atom$b #only retrieve atom b from trimmed file and assign to atombtrimmedfile
return(atombfromtrimmedfile)
}
#check if atomb function works 
#print(atomb("4AKE"))
#useplotb3 for each of the three files 
plotb3(atomb("4AKE"), typ="l", ylab="Bfactor")

plotb3(atomb("1AKE"), typ="l", ylab="Bfactor")

plotb3(atomb("1E4Y"), typ="l", ylab="Bfactor")
  

```














